---
title: "DistEdgeBurnLocs"
output: html_document
date: "2023-09-10"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries
```{r}
library(amt)
library(terra)
library(raster)
library(ggthemes)
library(scales)
library(janitor)
library(readxl)
library(sp)
library(lubridate)
library(writexl)
library(sf)
library(lme4)
theme_set(theme_minimal())
```

Importing points, and map of burned areas.
```{r}
#Importing points. Nesting birds have already been filtered out. Filtering for birds found in burned units (this attribute was added earlier in ArcGis)
summer_2022_burned_locs <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/Sum2022BurnedDist.xlsx") %>%
  clean_names() %>%
  filter(burned == 1,
         status != "D") 

#Importing map of burned areas, summer 2022.
burn_compartments_2022 <- raster::shapefile("/Volumes/Samsung_T5/BOBWHITE_DATA/Maps/Burn_Compartments2022/2022_BurnUnits.shp")

lcov <- raster::shapefile("/Volumes/Samsung_T5/BOBWHITE_DATA/Maps/SimpLcovLumped/SimplifiedLCOV_LumpedAbbreviated.shp")
```

Lets plot these and make sure everything is looking alright.
```{r}
#Plotting landcover
custom_palette_map <- c("AG" = "yellow", 
                    "HW" = "brown",
                    "NP" = "green",
                    "WL" = "blue",
                    "PP" = "darkgreen",
                    "SS" = "gray",
                    "UM" = "black")

plot(lcov, col = custom_palette_map[lcov$LCOV], main = "Landcover with Points")

#Adding points
points(summer_2022_burned_locs$easting, 
       summer_2022_burned_locs$northing, 
       col = summer_2022_burned_locs$band_numb, 
       pch = 20, 
       cex = 0.5)

#HERE IS AN ISSUE: Some of these points fall in AG, which was not actually burned (but within burn compartments). We must plot our points, extract covariates (lcov type) for each point, and filter out points in AG.

#First, lets double check all points are within burn compartments
#Let's plot to make sure...First we plot points (just a convenient way to set extent, since map of burn compartments is all of Ichauway)
plot(summer_2022_burned_locs$easting, 
       summer_2022_burned_locs$northing, 
       col = summer_2022_burned_locs$band_numb, 
       pch = 20, 
       cex = 0.5)

#Overlaying burn compartments. All points should disappear.
plot(burn_compartments_2022, col = as.factor(burn_compartments_2022$BURNDATE), add = TRUE)

#Adding points back in, to double check.
points(summer_2022_burned_locs$easting, 
       summer_2022_burned_locs$northing, 
       col = summer_2022_burned_locs$band_numb, 
       pch = 20, 
       cex = 0.5)

#All points do appear to fall in burned areas.
```

Okay, now we will try to plot our points onto our land cover map, extract lcov type for each point, and filter out AG.
First, we will try using the SP package. 
```{r}
#Setting projection and CRS to NAD 1983 16N
prj <- '+init=epsg:26916'
CRS('+init=epsg:26916')
proj4string(lcov) <- CRS("+init=epsg:26916") 


#Creating spatial points data frame.
summer_2022_burned_locs_SPDF <- SpatialPointsDataFrame(coordinates(cbind(summer_2022_burned_locs$easting, summer_2022_burned_locs$northing)), data = summer_2022_burned_locs, proj4string = CRS(prj))

#Verifying structure
str(summer_2022_burned_locs_SPDF)

#Plotting landcover, adding points. Not sure is this is necessary before running sp::over()...
plot(lcov, main = "Study site with Points") 
plot(summer_2022_burned_locs_SPDF, add = TRUE, col = "red", pch = 20)

#Double checking CRS is the same for SPDF and lcov before extracting covariates.
identical(proj4string(summer_2022_burned_locs_SPDF), proj4string(lcov))

#Extracting all lcov attributes
lcov_values <- sp::over(summer_2022_burned_locs_SPDF, lcov)

# Adding column on original data frame that specifies which lcov type each point falls into. Only adding LCOV attribute (there are a lot)
summer_2022_burned_locs$LCOV <- lcov_values$LCOV


#Viewing data frame
summer_2022_burned_locs %>%
  view()

#Counting LCOV types
summer_2022_burned_locs %>%
  count(LCOV, sort = TRUE)

#Let's plot to make sure this worked.
plot(lcov, col = custom_palette_map[lcov$LCOV], main = "Landcover with Points (SP)")

#Setting custom palette for points. We'll set AG to red for contrast.
custom_palette_points <- c("HW" = "brown",
                    "NP" = "green",
                    "PP" = "darkgreen",
                    "SS" = "gray",
                    "AG" = "red")

#This points should "disappear" if LCOV was extracted correctly. AG will be the exception and contrast with background.
points(x = summer_2022_burned_locs$easting, 
       y = summer_2022_burned_locs$northing, 
       col = custom_palette_points[summer_2022_burned_locs$LCOV], 
       pch = 20, 
       cex = 0.3)

#LCOV attribute appears to have been added accurately.


#Filtering out points located in AG. A lot of unnecessary columns. Let's refine and make it a tibble since its 2023.
summer_2022_burned_locs_clean <- summer_2022_burned_locs %>% 
  filter(LCOV != "AG") %>% 
  as_tibble() %>%
  dplyr::select(title, date_creat, date, northing, easting, band_numb, trap_locat, LCOV) %>%
  rename(date_created = date_creat)

#Counting LCOV to make sure there is no AG.
summer_2022_burned_locs_clean %>%
 count(LCOV, sort = TRUE)


#Looks like this worked. Points that fall in Ag have been filtered out and LCOV column is correct. This may be useful, later.
#Exporting this as an .xlsx
#write_xlsx(summer_2022_burned_locs_clean, "/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/summer_2022_burned_locs_clean.xlsx")
```

```{r}
#Reading burn compartments in again, but as an sf object
burn_compartments_2022 <- st_read("/Volumes/Samsung_T5/BOBWHITE_DATA/Maps/Burn_Compartments2022/2022_BurnUnits.shp")

# Converting burn_compartments_2022 polygons to line geometries that represent their boundaries. This is necessary to calc min distances later.
burn_compartment_boundaries <- st_boundary(burn_compartments_2022)

# Converting tibble of points to an sf object
summer_2022_burn_locs_clean_sf <- st_as_sf(summer_2022_burned_locs_clean, coords = c("easting", "northing"), crs = st_crs(burn_compartments_2022))

# Calculating the distance from each point to the nearest line boundary of the burn compartments, storing it in matrix
distance_matrix <- st_distance(summer_2022_burn_locs_clean_sf, burn_compartment_boundaries)

# Applying the function min() to each row (1 does this -- we'd use 2 for columns) in the distance matrix
min_distances <- apply(distance_matrix, 1, min)

# Add the minimum distance to the original tibble as a new column
summer_2022_burned_locs_clean$distance_to_edge <- min_distances

#Printing
print(summer_2022_burned_locs_clean)
```

Let's add a days since fire (DSF) column.
```{r}
summer_2022_burned_locs_clean <- summer_2022_burned_locs_clean %>%
  mutate(date = ymd(date))

burn_date_2022 <- ymd("2022-04-11")

summer_2022_burned_locs_clean$DSF <- as.numeric(summer_2022_burned_locs_clean$date - burn_date_2022)

summer_2022_burned_locs_clean %>%
  arrange(DSF)%>%
  view()
```

Let's try plotting DSF vs distance to edge with dist to edge data generated in R.
```{r}
summer_2022_burned_locs_clean %>%
  filter(date <= "2022-08-05") %>%
  ggplot(aes(DSF, distance_to_edge)) +
  geom_point() +
  geom_smooth() +
  labs(x = "DSF",
       y = "Distance to edge",
       title = "Distance to edge over time",
       subtitle = "Only looking at NOBO located in burn compartments")
```

Not the relationship we were hoping to see. Let's try with distance to edge data that was generated in ArcMap
```{r}
summer_2022_burned_locs_arc <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/summer_2022_burn_locs_wDTE_arc.xlsx") %>%
  mutate(date = as.Date(date)) %>%
  filter(date <= "2022-08-06")

head(summer_2022_burned_locs_arc)

summer_2022_burned_locs_arc$DSF <- as.numeric(summer_2022_burned_locs_arc$date - burn_date_2022) 

summer_2022_burned_locs_arc %>%
  ggplot(aes(DSF, NEAR_DIST)) +
  geom_point() +
  geom_smooth() +
  labs(x = "DSF",
       y = "Distance to edge",
       title = "Distance to edge of burn compartments following prescribed fire")

summer_2022_burned_locs_arc %>%
  ggplot(aes(NEAR_DIST)) +
  geom_histogram(binwidth = 10) +
  labs(x = "Distance to edge (m)",
       y = "",
       title = "Histogram: Distance to edge of burn compartments",
       subtitle = "Summer 2022")

#Importing "fate" spreadsheet, which has sex and age-class associated with each band number (unlike daily locations data)
fate <- rio::import("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/fate_clean_2021_22.xlsx", setclass = "tibble") %>%
  clean_names() %>%
  rename(band_numb = band_number)

#Adding sex and age to each observation/recorded location.
summer_2022_burned_locs_arc_age_sex <- summer_2022_burned_locs_arc %>%
  left_join(fate %>% 
              dplyr::select(band_numb, age, sex), by = "band_numb")

summer_2022_burned_locs_arc_age_sex %>%
  arrange(band_numb) %>%
  view()

summer_burned_locs_female <- summer_2022_burned_locs_arc_age_sex %>%
  filter(sex == "F")

summer_burned_locs_female %>%
  ggplot(aes(DSF, NEAR_DIST)) +
  geom_point() +
  geom_smooth() +
  labs(x = "DSF",
       y = "Distance to edge",
       title = "Hens: distance to edge of burn compartments following prescribed fire")

summer_burned_locs_male <- summer_2022_burned_locs_arc_age_sex %>%
  filter(sex == "M")

summer_burned_locs_male %>%
  ggplot(aes(DSF, NEAR_DIST)) +
  geom_point() +
  geom_smooth() +
  labs(x = "DSF",
       y = "Distance to edge",
       title = "Males: distance to edge of burn compartments following prescribed fire")


summer_burned_locs_adults <- summer_2022_burned_locs_arc_age_sex %>%
  filter(age == "A")

summer_burned_locs_adults %>%
  ggplot(aes(DSF, NEAR_DIST)) +
  geom_point() +
  geom_smooth() +
  labs(x = "DSF",
       y = "Distance to edge",
       title = "Adults: distance to edge of burn compartments following prescribed fire")

summer_burned_locs_juvs <- summer_2022_burned_locs_arc_age_sex %>%
  filter(age == "J")

summer_burned_locs_juvs %>%
  ggplot(aes(DSF, NEAR_DIST)) +
  geom_point() +
  geom_smooth() +
  labs(x = "DSF",
       y = "Distance to edge",
       title = "Juveniles: distance to edge of burn compartments following prescribed fire")
```

Lets run a mixed model and see if probability of locating a bobwhite in a burned area depends on DSF
First, we must extract lcov for each point, and set any points in AG to "unburned".
```{r}
#Importing locations with "burned" column (0 = unburned loc, 1 = burned loc), filtering out dead birds and for proper time frame.
summer_2022_locs_burned_and_unburned <- read_xlsx("/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/Sum2022BurnedDist.xlsx") %>%
  clean_names() %>%
  filter(status != "D") %>%
  mutate(date = as.Date(date)) %>%
  filter(date <= "2022-08-05",
         date >= "2022-04-11")

#Counting burned and unburned locations.
summer_2022_locs_burned_and_unburned %>% 
  count(burned)

#Once again, we need to plot points and add LCOV to filter out AG.
#Setting projections
prj <- '+init=epsg:26916'
CRS('+init=epsg:26916')
proj4string(lcov) <- CRS("+init=epsg:26916") 


#Creating spatial points data frame of burned and unburned points.
summer_2022_locs_burned_unburned_SPDF <- SpatialPointsDataFrame(coordinates(cbind(summer_2022_locs_burned_and_unburned$easting, summer_2022_locs_burned_and_unburned$northing)), data = summer_2022_locs_burned_and_unburned, proj4string = CRS(prj))

#Verifying structure
str(summer_2022_locs_burned_unburned_SPDF)

#Plotting landcover, adding points. Not sure is this is necessary before running sp::over()...
plot(lcov, main = "Study site with Points") 
plot(summer_2022_locs_burned_unburned_SPDF, add = TRUE, col = "red", pch = 20)

#Double checking CRS is the same for SPDF and lcov before extracting covariates.
identical(proj4string(summer_2022_locs_burned_unburned_SPDF), proj4string(lcov))

#Extracting all lcov attributes
lcov_values_burned_unburned <- sp::over(summer_2022_locs_burned_unburned_SPDF, lcov)

# Adding column on original data frame that specifies which lcov type each point falls into. Only adding LCOV attribute (there are a lot)
summer_2022_locs_burned_and_unburned$LCOV <- lcov_values_burned_unburned$LCOV


#Viewing data frame
summer_2022_locs_burned_and_unburned %>%
  view()

#Counting LCOV types
summer_2022_locs_burned_and_unburned %>%
  count(LCOV, sort = TRUE)

#Let's plot to make sure this worked.
plot(lcov, col = custom_palette_map[lcov$LCOV], main = "Landcover with Points")

#Setting custom palette for points. We'll set AG to red for contrast.
custom_palette_points <- c("HW" = "brown",
                    "NP" = "green",
                    "PP" = "darkgreen",
                    "SS" = "gray",
                    "AG" = "yellow")

#This points should "disappear" if LCOV was extracted correctly. 
points(x = summer_2022_locs_burned_and_unburned$easting, 
       y = summer_2022_locs_burned_and_unburned$northing, 
       col = custom_palette_points[summer_2022_locs_burned_and_unburned$LCOV], 
       pch = 20, 
       cex = 0.1)
#This looks good

#A lot of unnecessary columns. Let's refine and make it a tibble since.
summer_2022_locs_burned_and_unburned_clean <- summer_2022_locs_burned_and_unburned %>% 
  as_tibble() %>%
  dplyr::select(title, date_creat, date, northing, easting, band_numb, trap_locat, burned, LCOV) %>%
  rename(date_created = date_creat)

#Quick look
head(summer_2022_locs_burned_and_unburned_clean)

#Counting
summer_2022_locs_burned_and_unburned_clean %>% 
  count(burned, LCOV, sort = T)

#writing an expression so all points that fall into AG are unburned.
summer_2022_locs_burned_and_unburned_clean <- summer_2022_locs_burned_and_unburned_clean %>%
  mutate(burned = ifelse(LCOV == "AG", 0, burned))

#Counting again to make sure that worked.
#Looks good.
summer_2022_locs_burned_and_unburned_clean %>% 
  count(burned, LCOV, sort = T)

```

Adding a days since fire (DSF column), running mixed model, and viewing summary.
```{r}
summer_2022_locs_burned_and_unburned_clean$DSF = as.numeric(summer_2022_locs_burned_and_unburned_clean$date - burn_date_2022)

summer_2022_locs_burned_and_unburned_clean %>%
  arrange(DSF) %>%
  view()

#write_xlsx(summer_2022_locs_burned_and_unburned_clean, "/Volumes/Samsung_T5/BOBWHITE_DATA/Clean/summer_2022_locs_burned_and_unburned_clean.xlsx")

#Fitting glm
prob_burned_area <- glmer(burned ~ DSF + (1|band_numb),
                 data = summer_2022_locs_burned_and_unburned_clean,
                 na.action = "na.pass",
                 family = "binomial")
 
#Viewing summary 
summary(prob_burned_area) 
```

